---
title: "raceranalysis"
author: "shrey agrawal"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

#setup and cleaning

Importing and cleaning data, removing an outlier
```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
require(mosaic)

##Importing Data
allData = read.csv("https://stat2games.sites.grinnell.edu/data/racekartb/getdata.php")

#cleaning
NewDataRestrictedCars = filter(allData, Game > 633, FinishedTime > 0.1, CarID == 1|CarID == 15 |CarID==29)
NewDataUnrestricted = filter(allData, Game > 633, FinishedTime > 0.1)
  
NewDataRestrictedCars$GroupID <- tolower(NewDataRestrictedCars$GroupID)
NewDataRestrictedCars$PlayerID <- tolower(NewDataRestrictedCars$PlayerID)
NewDataUnrestricted$GroupID <- tolower(NewDataUnrestricted$GroupID)
NewDataUnrestricted$PlayerID <- tolower(NewDataUnrestricted$PlayerID)


#REMOVING AN OUTLIER
NewDataRestrictedCars = filter(NewDataRestrictedCars, FinishedTime != 99.90)
NewDataUnrestricted = filter(NewDataUnrestricted, FinishedTime != 99.90)

names(NewDataUnrestricted)
str(NewDataUnrestricted$CarID)
```


#In restricted, which is the fastest car
```{r}
NewDataRestrictedCars = arrange(NewDataRestrictedCars, FinishedTime)
print("TABLE SORTED BY FINISHED TIME")
head(NewDataRestrictedCars)

NewDataRestrictedCars = arrange(NewDataRestrictedCars, TopSpeed)
print("TABLE SORTED BY TOPSPEED TIME")
head(NewDataRestrictedCars)

NewDataRestrictedCars = arrange(NewDataRestrictedCars, TimeTo30)
print("TABLE SORTED BY TIMETO30 TIME")
head(NewDataRestrictedCars)

restrictedplot1 = ggplot(data = NewDataRestrictedCars, aes(x = CarID, y = FinishedTime)) + geom_point()  + aes(colour = Track) + theme(legend.position = "right") + labs(title = "Restricted to 1,15,29. FinishedTime vs CarID, colored by Track")
restrictedplot1

restrictedplot2 = ggplot(data = NewDataRestrictedCars, aes(x = CarID, y = TopSpeed)) + geom_point()  + aes(colour = Track) + theme(legend.position = "right") + labs(title = "Restricted to 1,15,29. TopSpeed vs CarID, colored by Track")
restrictedplot2

restrictedplot3 = ggplot(data = NewDataRestrictedCars, aes(x = CarID, y = TimeTo30)) + geom_point()  + aes(colour = Track) + theme(legend.position = "right") + labs(title = "Restricted to 1,15,29. TimeTo30 vs CarID, colored by Track")
restrictedplot3

```



#all cars straight track - 3PLOTS, 1TABLE HERE
```{r}

straightData = filter(NewDataUnrestricted, NewDataUnrestricted$Track == "StraightTrack")
straightData = arrange(straightData, FinishedTime)
head(straightData)

plotfinishvscarstraight = ggplot(data = straightData, aes(x = CarID, y = FinishedTime)) + geom_point()  + 
  aes(colour = PlayerID) + 
  theme(legend.position = "right") + 
  labs(title = "StraightTrack results only. FinishTime vs CarID, colored by PlayerID")
plotfinishvscarstraight

plottopspeedvscarstraight = ggplot(data = straightData, aes(x = CarID, y = TopSpeed)) + geom_point()  + 
  aes(colour = PlayerID) + 
  theme(legend.position = "right") + 
  labs(title = "StraightTrack results only. TopSpeed vs CarID, colored by PlayerID")
plottopspeedvscarstraight

plottimeto30vscarstraight = ggplot(data = straightData, aes(x = CarID, y = TimeTo30)) + geom_point()  + 
  aes(colour = PlayerID) + 
  theme(legend.position = "right") + 
  labs(title = "StraightTrack results only. Timeto30 vs CarID, colored by PlayerID")
plottimeto30vscarstraight

```


#all cars oval track - 3PLOTS, 1TABLE HERE
```{r}

ovalData = filter(NewDataUnrestricted, NewDataUnrestricted$Track == "OvalTrack")
ovalData = arrange(ovalData, FinishedTime)
head(ovalData)

plotfinishvscaroval = ggplot(data = ovalData, aes(x = CarID, y = FinishedTime)) + geom_point()  + 
  aes(colour = PlayerID) + 
  theme(legend.position = "right") + 
  labs(title = "OvalTrack results only. FinishTime vs CarID, colored by PlayerID")
plotfinishvscaroval

plottopspeedvscaroval = ggplot(data = ovalData, aes(x = CarID, y = TopSpeed)) + geom_point()  + 
  aes(colour = PlayerID) + 
  theme(legend.position = "right") + 
  labs(title = "OvalTrack results only. TopSpeed vs CarID, colored by PlayerID")
plottopspeedvscaroval

plottimeto30vscaroval = ggplot(data = ovalData, aes(x = CarID, y = TimeTo30)) + geom_point()  + 
  aes(colour = PlayerID) + 
  theme(legend.position = "right") + 
  labs(title = "OvalTrack results only. Timeto30 vs CarID, colored by PlayerID")
plottimeto30vscaroval

```



#q1. who was the fastest on the oval track?
```{r}

plot1 = ggplot(data = NewDataUnrestricted, aes(x = Track, y = FinishedTime)) + geom_point()  + aes(colour = as.factor(PlayerID)) + theme(legend.position = "right") + labs(title = "Track vs FinishTime Colored by PlayerID")

plot1
```


#q2. Which tire, body, engine combination is fastest on the STRAIGHT track? LOWEST FINISH TIME
```{r}

straightData = filter(NewDataUnrestricted, NewDataUnrestricted$Track == "StraightTrack")
straightData = arrange(straightData, FinishedTime)
head(straightData)

plot2 = ggplot(data = straightData, aes(x = Engine, y = FinishedTime)) + geom_point()  + 
  aes(colour = Body) + 
  facet_grid(~Tire) + 
  theme(legend.position = "right") + 
  labs(title = "StraightTrack results only. FinishTime vs Engine faceted by Tire, colored by Body")

plot2

```


#q3. Which tire, body, engine combination is fastest on the OVAL track? LOWEST FINISH TIME
```{r}

ovalData = filter(NewDataUnrestricted, NewDataUnrestricted$Track == "OvalTrack")
ovalData = arrange(ovalData, FinishedTime)
head(ovalData)

plot3 = ggplot(data = ovalData, aes(x = Engine, y = FinishedTime)) + geom_point()  + aes(colour = Body) + facet_wrap(~Tire, ncol = 4) + theme(legend.position = "right") + labs(title = "OvalTrack results only. FinishTime vs Engine faceted by Tire, colored by Body")

plot3
```



#q4. Which tire, body, engine combination has the highest TOP SPEED REACHED
```{r}
plot4 = ggplot(data = NewDataUnrestricted, aes(x = Engine, y = TopSpeed)) + 
  geom_point()  + 
  aes(colour = Body) + 
  facet_wrap(~Tire, ncol = 4) + 
  theme(legend.position = "right") + labs(title = "TopSpeed vs Engine faceted by Tire, colored by Body")

plot4

```



#Some extra analysis in table form with Mean Finish TIMES.
```{r}
group1 = group_by(NewDataUnrestricted, Track, Engine, Tire, Body) 
group2 = summarise(group1, 
        TimesPlayed = n(),
        MeanFinishedTime = mean(FinishedTime, na.rm = TRUE), 
        SDFinishTime = sd(FinishedTime, na.rm= TRUE))

group2=group2[order(group2$MeanFinishedTime),]
group2

```



#anova test
```{r}
anovatest1 <- aov(FinishedTime ~ Track*Engine*Tire*Body, data = NewDataUnrestricted)
summary(anovatest1)

anovatest1 <- aov(FinishedTime ~ Engine*Tire*Body, data = ovalData)
summary(anovatest1)

anovatest1 <- aov(FinishedTime ~ Engine*Tire*Body, data = straightData)
summary(anovatest1)

anovatest2 <- aov(FinishedTime ~ Tire, data = straightData)
summary(anovatest2)

anovatest3 <- aov(FinishedTime ~ Body, data = straightData)
summary(anovatest3)

anovatest3 <- aov(FinishedTime ~ Engine, data = straightData)
summary(anovatest3)

#ANOVA FOR CARID
anovatestcar <- aov(FinishedTime ~ CarID, data = NewDataUnrestricted)
summary(anovatestcar)
```








