---
title: "Untitled"
author: "kuiper"
date: '2022-07-15'
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## initial setup
This online tutorial is designed to introduce you to using R and RStudio to make the graphs using the NYPD. In this tutorial, you can edit and execute R code and see the results within your browser. You can also learn more details about coding in R using our RTutorials from our Stat2Labs page {{include linkhttps://stat2labs.sites.grinnell.edu/RTutorials.html}}


 


In R, text and coding are often separated. The places where we type our code are called Code Chunks.





```{r}
library(readr)
NYPDdf <- read_csv("C:/Users/kuipers/Dropbox/MAP_2022Summer/Analytics/NYPD/NYPD_BarCharts/NYPDBarData.csv")
dim(NYPDdf)
names(NYPDdf)
```

## Making plots 1 :

Make plots from page 1{{https://dataspace.sites.grinnell.edu/nypd1.html}}: USE ONLY MOSAIC AND GGFORMULA packages for this


```{r pressure, echo=FALSE}

```


Throughout this tutorial we are using an R package called mosaic {{indlude link}} that is designed to for people new to computer programming.
