---
title: "NYPD page 4"
author: "shrey agrawal"
date: '2022-07-18'
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## initial setup
This online tutorial is designed to introduce you to using R and RStudio to make the graphs using the NYPD. In this tutorial, you can edit and execute R code and see the results within your browser. You can also learn more details about coding in R using our RTutorials from our Stat2Labs page {{include linkhttps://stat2labs.sites.grinnell.edu/RTutorials.html}}

In R, text and coding are often separated. The places where we type our code are called Code Chunks.

```{r}
library(readr)
library(learnr)
NYPDdf <- read_csv("NYPD_BarCharts/NYPDBarData.csv")
dim(NYPDdf)
names(NYPDdf)
```

```{r setup, include=FALSE}
#setup
NYPDdf <- mutate(NYPDdf, CrimeType = replace(CrimeType, CrimeType == "Tresspass" ,"Trespass"))

####### Custom Colors #######
customColors <- c("#a6cee3", "#1f78b4", "#b2df84", "#33a02c",
                  "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00")

####### Force Variables #######
ForceVariables <- c(
                    "HandCuff" = "HandCuff",
                    "Firearm" = "Firearm",
                    "CEW" = "CEW",
                    "PepperSpray" = "PepperSpray",
                    "Other" = "Other",
                    "Verbal" = "Verbal",
                    "Summons" = "Summons"
                    )

####### Race Variables #######
RaceVariables <- c("Black" = "Black",
                   "Hispanic" = "Hispanic",
                   "White" = "White",
                   "Other" = "Other")

####### Gender Variables #######
GenderVariables <- c("Female" = "Female",
                     "Male" = "Male",
                     "Other" = "Other")

####### Crime Type Variables #######
CrimeVariables <- c("Assault" = "Assault",
                    "Trespass" = "Trespass",
                    "Weapon" = "Weapon",
                    "Theft" = "Theft",
                    "Substance" = "Substance",
                    "Other" = "Other")

####### Categorical Variables #######
CatOptions = c("Race" = "Race",
               "Gender" = "Gender",
               "Crime Type" = "CrimeType",
               "Year" = "Year")

####### Quantitative Variables #######
QuanOptions = c("Stopped" = "Stopped",
                "Frisked" = "Frisked",
                "Searched" = "Searched",
                "Arrested" = "Arrested",
                ForceVariables)

```
## Making plots 1 :

Make plots from page 1{{https://dataspace.sites.grinnell.edu/nypd1.html}}: USE ONLY MOSAIC AND GGFORMULA packages for this


```{r pressure, echo=FALSE}


```


Throughout this tutorial we are using an R package called mosaic {{indlude link}} that is designed to for people new to computer programming.
